---
title: "Big Data In Finance Individual Assignment"
author: "Louise Fallon"
output: pdf_document
---

```{r setup, include=FALSE}
library(readxl) #for read_excel
library(ggplot2) #for ggplot
library(forecast) #for ggseasonplot and autoplot etc
library(reshape2) #for melt
library(condformat) #for condformat and rule_fill_discrete
library(glmnet)
knitr::opts_chunk$set(cache = TRUE, echo=FALSE)
```

## Data

```{r}
#load data
returns <- as.data.frame(read_excel("Data.xlsx", sheet = 1, col_names = TRUE, skip=3))
#remove row of NAs at the bottom
returns <- returns[complete.cases(returns),]
#add date column header
names(returns)[1] <- "date"
#plot boxplot of returns
meltedreturns <- melt(returns, id.vars = "date")
ggplot(meltedreturns, aes(x=variable, y=value)) + geom_boxplot(col="#000066") + theme_bw() + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust=0.5)) + xlab("Industry") + ylab("Daily Returns")
```

Industry returns are generally centered around 0, the lowest average return is for Coal at `r format(min(colMeans(returns[,2:50])),digits=3)`, and the highest is for Gold at `r format(max(colMeans(returns[,2:50])),digits=3)`, with variance ranging from 
`r format(min(apply(returns[,2:50], 2, var)),digits=3)` for Household, and `r format(max(apply(returns[,2:50], 2, var)),digits=3)` for Coal.

To use this for prediction we are looking for a relationship between industry returns and either contemporaneous or lagged returns in the same industry, and in other industries. If industry momentum is a real phenomenon, then we would expect to see a correlation between current returns with previous returns, within an industry

To take a naive initial look at whether this type of pattern is found in the data, we inspect the correlations between the variables and their 1st and 2nd lags.

```{r}
##create correlation matrix
corrmatrix <- round(cor(returns[,2:50]), 2)
melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "Contemporaneous Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[2:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-1),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-1") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[3:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-2),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "Twice Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-2") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[4:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-3),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "3-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-3") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[5:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-4),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "4-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-4") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```


```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[6:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-5),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "5-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-5") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```


```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[7:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-6),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "6-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-6") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[8:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-7),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "7-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-7") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[9:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-8),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "8-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-8") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```


Visual Check that the variables seem to be correlated:

```{r}
meltedreturns$actualdate <- as.Date(as.character(meltedreturns$date), '%Y%m%d')
meltedreturns.agrifood <- meltedreturns[meltedreturns$variable  %in% c("Agric","Beer "),]
ggplot(meltedreturns.agrifood, aes(x=actualdate, y=value, col=variable)) + geom_line() + theme_bw() + theme( axis.text.x = element_text(angle = 90, vjust=0.5)) + ylab("Daily Returns")
```


```{r}
meltedreturns$actualdate <- as.Date(as.character(meltedreturns$date), '%Y%m%d')
meltedreturns.agrifood <- meltedreturns[meltedreturns$variable  %in% c("Oil  ","ElcEq"),]
ggplot(meltedreturns.agrifood, aes(x=actualdate, y=value, col=variable)) + geom_line() + theme_bw() + theme( axis.text.x = element_text(angle = 90, vjust=0.5)) + ylab("Daily Returns")
```

Possible explanation is that these are "business cycle" trends so all seem to be correlated contemporaneously.

```{r}
##create correlation matrix
for (i in 1:49){
  for (j in 1:49){
#correlating the current value of i, with the lagged value of j
#starting from second period so that lagged values can be calculated
#the +1 is to avoid the date column
    corrmatrix[i,j] <- round(cor(returns[13:nrow(returns),i+1],
                                 returns[1:(nrow(returns)-12),j+1]), 2)
  }
}

melteddf<- melt(corrmatrix)

ggplot(data = melteddf, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    labs(x = "", y = "", title = "12-Lagged Correlation Matrix") +
    theme_bw() + ylab("Industry returns at time t") + xlab("Industry returns at time t-12") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6),
          axis.text.y = element_text(size=6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_blank()) + 
  scale_fill_gradient2(low = "#ce1254",mid="#ffffff", high = "#006699",
                       midpoint = 0, limit = c(-.2,.2), space = "Lab",
                       name="Pearson\nCorrelation") 
```


Check for autocorrelation within time series
```{r eval=FALSE}
Agricts <- ts(returns[,2], start=c(1,20), frequency=252)
Foodts <- ts(returns[,3], start=c(1,20), frequency=252)
Sodats <- ts(returns[,4], start=c(1,20), frequency=252)
Beerts <- ts(returns[,5], start=c(1,20), frequency=252)
Goldts <- ts(returns[,28], start=c(1,20), frequency=252)

fit = stl(Agricts, s.window='periodic')
plot(fit)
#season plot showing no clear seasonality
ggseasonplot(Agricts, year.labels=TRUE, year.labels.left=TRUE)
#lag plot showing no clear lag correlation
gglagplot(Agricts)
##ACF showing that effect on shocks seems to reduce, and that there are some lags that seem significant 
ggAcf(Agricts) + theme_bw() + theme(panel.grid.minor=element_blank()) + scale_x_continuous(breaks=c(2,21,40,42,75,77,252))
ggAcf(Foodts) + theme_bw() + theme(panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks=c(4,5,15,28,49,53,252))
ggAcf(Sodats) + theme_bw() + theme(panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks=c(86,166))
ggAcf(Beerts) + theme_bw() + theme(panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks=c(90))
ggAcf(Goldts) + theme_bw() + theme(panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks=c(7,14,48,64,78,252,264))
#no evidence of AC for lag =1
Box.test(Agricts, lag=1)
#some for lag=2
Box.test(Agricts, lag=2)
#some for lag=3
Box.test(Agricts, lag=3)
#none for lag=4
Box.test(Agricts, lag=4)
#none for lag=5
Box.test(Agricts, lag=5)
```

#Tests for Auto-correlation with lagged same-industry variables.

Using a Box-Pierce test each variable was tested for autocorrelation with lags up to 10 working days. The results displayed below show whether each of these are statistically significant at the 10% level, indicating that there is evidence of autocorrelation.

```{r cache=FALSE}
boxpiercevalue <- matrix(0, nrow = 49, ncol = 10) 
for (i in 1:49){
  for (j in 1:10){
    boxpiercevalue[i,j] <- round(Box.test(ts(returns[,i+1],
                                       start=c(1,20),
                                       frequency=252), lag=j)$p.value,3)
  }
}

boxpiercevaluedf <- as.data.frame(boxpiercevalue)
boxpiercevaluedf <- cbind(names(returns)[2:50],boxpiercevaluedf)
names(boxpiercevaluedf) <- c("Industry","Lag 1","Lag 2","Lag 3",
                        "Lag 4","Lag 5","Lag 6","Lag 7",
                        "Lag 8","Lag 9","Lag 10")

pval <- 0.1
condtable <- condformat(boxpiercevaluedf)
condtable <- condtable + rule_fill_discrete(2,
                     expression = boxpiercevaluedf[,2] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(3,
                     expression = boxpiercevaluedf[,3] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(4,
                     expression = boxpiercevaluedf[,4] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(5,
                     expression = boxpiercevaluedf[,5] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(6,
                     expression = boxpiercevaluedf[,6] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(7,
                     expression = boxpiercevaluedf[,7] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(8,
                     expression = boxpiercevaluedf[,8] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(9,
                     expression = boxpiercevaluedf[,9] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(10,
                     expression = boxpiercevaluedf[,10] < pval,
                     colours = c("TRUE" = "#c2f0c2")) +
                      rule_fill_discrete(11,
                     expression =  boxpiercevaluedf[,11] < pval,
                     colours = c("TRUE" = "#c2f0c2"))
condtable
```

#Model Building

```{r eval=FALSE, results="hide"}
df <- returns
#when adding in lags need to take away the last row bc it doesn't have a y
windowlength <- 80

historicalmeanpredictions <- matrix(0, ncol = 49, nrow = (nrow(df)+1))
linearpredictions <- matrix(0, ncol = 49, nrow = (nrow(df)+1))
lassopredictions <- matrix(0, ncol = 49, nrow = (nrow(df)+1))

linearcoefs <- matrix(0, ncol = (ncol(df)-1), nrow = (nrow(df)+1)) 
lassocoefs <- matrix(0, ncol = (ncol(df)-1), nrow = (nrow(df)+1)) 

#windows run from the start of the window to the last full record
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), 2:ncol(df)]
        #take away target column from predictors
        train.predictors <- train.predictors[, -j]
        #combine for one df
        train.full <- cbind(train.target,train.predictors)
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, 2:ncol(df)]
        tau.newdata <- train.newdata[, -j]
        
        #find historical mean value
        historicalmeanpredictions[tau+1,j] <- mean(train.target)

        #find linear model prediction
        temp.mdl.linear <- lm(train.target ~ ., data=train.full)
        linearpredictions[tau+1,j] <- predict(temp.mdl.linear,
                                        newdata= tau.newdata)
        linearcoefs[tau+1,] <- coef(temp.mdl.linear)
        
        #find cv with provided vals lasso prediction
        temp.mdl.lasso.mancv <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5,
                               lambda = c(0.01,0.025,0.05))
        lassopredictions[tau+1,j] <- predict(temp.mdl.lasso.mancv,
                                        newx = as.matrix(tau.newdata),
                                    type = "response")
        lassocoefs[tau+1,] <- as.vector(coef(temp.mdl.lasso.mancv))
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
```

```{r}
#Out of Sample Accuracy
actualsdf <- returns[(windowlength+1):(nrow(returns)),2:ncol(returns)]

historicalmeanpredictions.clean <- as.data.frame(historicalmeanpredictions)[(windowlength+1):(nrow(historicalmeanpredictions)-1),]
historicalmeanSE <- (historicalmeanpredictions.clean - actualsdf)^2
historicalmeanSSE <- colSums(historicalmeanSE)

linear.clean <- as.data.frame(linearpredictions)[(windowlength+1):(nrow(linearpredictions)-1),]
linearSE <- (linear.clean - actualsdf)^2
linearSSE <- colSums(linearSE)

lasso.clean <- as.data.frame(lassopredictions)[(windowlength+1):(nrow(lassopredictions)-1),]
lassoSE <- (lasso.clean - actualsdf)^2
lassoSSE <- colSums(lassoSE)

R2oos <- data.frame(linear = round(1-(linearSSE/historicalmeanSSE),3),
                    lasso =  round(1-(lassoSSE/historicalmeanSSE),3))
R2oos
```

```{r}
#For Goyel plots

#function to get cumulative sums of a dataframe
colCumSums <- function(x) {
  for(i in seq_len(dim(x)[2])) { x[,i] <- cumsum(x[,i]) }; x
}

```



## References
https://www.treasury.gov/resource-center/data-chart-center/interest-rates/Pages/TextView.aspx?data=billRatesYear&year=2015

```{r eval=FALSE}
oldlassocoefs <- matrix(0, ncol = (ncol(df)-1), nrow = (nrow(df)+1)) 
manlassocoefs <- matrix(0, ncol = (ncol(df)-1), nrow = (nrow(df)+1)) 

        #find CV lasso model prediction
        temp.mdl.lasso <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5)
        predictions[tau+1,3] <- predict(temp.mdl.lasso,
                                        newx= as.matrix(df[tau,3:ncol(df)]))
        oldlassocoefs[tau+1,] <- as.vector(coef(temp.mdl.lasso))
        
        #find manual lasso model prediction
        temp.mdl.lasso.man <- glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1)
        predictions[tau+1,4] <- predict(temp.mdl.lasso.man, newx = as.matrix(df[tau,3:ncol(df)]),
                                    type = "response",
                                    s = 0.03)
        manlassocoefs[tau+1,] <- as.vector(coef(temp.mdl.lasso.man, s=0.03))
```

