This is in RMD format for the benefit of using chunks, but is not a final report.
Many of these chunks contain models that take many minutes to run.

Each of the outputs of the large models is saved into an .RData file, all of which have been included in the submission, together with all outputs that are required for the main report.

Cross-validation is done regularly (randomly) within window, which should not invalidate the time-relevant property of the full approach, because at any time tau, the window tau-80:tau will be known to the user/model, and so they will be able to perform regular cross-validation within that window, and train on all the data in that window, without any time limitations to predict the tau+1th return. 

It is noted that the implementation of this rolling window and multiple re-training approach would potentially run faster using a sklearn Python implementation, but R was used for this project for educational purposes, as Python was used for the group project.

```{R}
library(glmnet) #for cv.glmnet (LASSO and elastic nets)
library(randomForest) #for randomForest
library(gbm) #for gbm
library(condformat) #for condformat
library(readxl) #for condformat
```

```{r loaddata}
#load data
returns <- as.data.frame(read_excel("Data.xlsx", sheet = 1, col_names = TRUE, skip=3))
#remove lagged spaces in column names
for (i in 1:ncol(returns)) colnames(returns)[i] <- gsub(" ","",colnames(returns)[i])
#remove row of NAs at the bottom
returns <- returns[complete.cases(returns),]
#add date column header
names(returns)[1] <- "date"
```

```{r overallparams}
windowlength <- 80
```

```{r 1lagprep}
df.1 <- returns
```

```{r own1lagmodel}
set.seed(88)
historicalmeanpredictions <- matrix(0, ncol = 49, nrow = (nrow(df.1)+1))
linearpredictions.1.own <- matrix(0, ncol = 49, nrow = (nrow(df.1)+1))

df <- df.1
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #chose only the own lag variable
        train.predictors <- as.data.frame(df[(tau-windowlength+1):(tau-1), j+1])
        colnames(train.predictors) <- colnames(df)[j+1]

        #combine for one df
        train.full <- (cbind(train.target,train.predictors))
        
        #newdata for predicting tau + 1
        tau.newdata <- as.data.frame(df[tau, j+1])
        colnames(tau.newdata) <- colnames(df)[j+1]
        
        #find historical mean value
        historicalmeanpredictions[tau+1,j] <- mean(train.target)

        #find linear model prediction
        temp.mdl.linear <- lm(train.target ~ ., data=train.full)
        linearpredictions.1.own[tau+1,j] <- predict(temp.mdl.linear,
                                        newdata= tau.newdata)
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(historicalmeanpredictions,
     linearpredictions.1.own,
     file="linearownpredictions.RData")
```

```{r alllag1models}
set.seed(88)
df <- df.1

linearpredictions.1.all <- matrix(0, ncol = 49, nrow = (nrow(df.1)+1))
lassopredictionscv.1.all <- matrix(0, ncol = 49, nrow = (nrow(df.1)+1))

#windows run from the start of the window to the last full record
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), 2:ncol(df)]

        #combine for one df
        train.full <- cbind(train.target,train.predictors)
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, (2:ncol(df))]

        #find linear model prediction
        temp.mdl.linear <- lm(train.target ~ ., data=train.full)
        linearpredictions.1.all[tau+1,j] <- predict(temp.mdl.linear,
                                        newdata= tau.newdata)

        #find cv
        temp.mdl.lasso.allcv <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5)
        lassopredictionscv.1.all[tau+1,j] <- predict(temp.mdl.lasso.allcv,
                                        newx = as.matrix(tau.newdata),
                                    type = "response")
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(linearpredictions.1.all,
     lassopredictionscv.1.all,
     file="lag1allpredictions.RData")
```

```{r lag1performanceprep}
actualsdf <- returns[(windowlength+1):(nrow(returns)),2:ncol(returns)]

historicalmeanpredictions.clean <- as.data.frame(historicalmeanpredictions)[(windowlength+1):(nrow(historicalmeanpredictions)-1),]
historicalmeanSE <- (historicalmeanpredictions.clean - actualsdf)^2
historicalmeanSSE <- colSums(historicalmeanSE)

linearpredictions.1.own.clean <- as.data.frame(linearpredictions.1.own)[(windowlength+1):(nrow(linearpredictions.1.own)-1),]
linear.1.ownSE <- (linearpredictions.1.own.clean - actualsdf)^2
linear.1.ownSSE <- colSums(linear.1.ownSE)

linearpredictions.1.all.clean <- as.data.frame(linearpredictions.1.all)[(windowlength+1):(nrow(linearpredictions.1.all)-1),]
linear.1.allSE <- (linearpredictions.1.all.clean - actualsdf)^2
linear.1.allSSE <- colSums(linear.1.allSE)

lassopredictionscv.1.all.clean <- as.data.frame(lassopredictionscv.1.all)[(windowlength+1):(nrow(lassopredictionscv.1.all)-1),]
lasso.cv.1.allSE <- (lassopredictionscv.1.all.clean - actualsdf)^2
lasso.cv.1.allSSE <- colSums(lasso.cv.1.allSE)
```

```{r 4lagprep}
lag.1 <- returns[4:nrow(returns),] 
for (i in 2:ncol(lag.1)) colnames(lag.1)[i] <- paste(colnames(lag.1)[i],"_lag_1",sep="") 
lag.2 <- returns[3:(nrow(returns)-1),2:ncol(returns)]
for (i in 1:ncol(lag.2)) colnames(lag.2)[i] <- paste(colnames(lag.2)[i],"_lag_2",sep="") 
lag.3 <- returns[2:(nrow(returns)-2),2:ncol(returns)]
for (i in 1:ncol(lag.3)) colnames(lag.3)[i] <- paste(colnames(lag.3)[i],"_lag_3",sep="")
lag.4 <- returns[1:(nrow(returns)-3),2:ncol(returns)]
for (i in 1:ncol(lag.4)) colnames(lag.4)[i] <- paste(colnames(lag.4)[i],"_lag_4",sep="")
df.4 <- cbind(lag.1,lag.2,lag.3,lag.4)
```

```{r own4lagmodel}
set.seed(88)
historicalmeanpredictions.4 <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
linearpredictions.4.own <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
lassopredictionscv.4.own <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
df <- df.4
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), c(j+1,j+50,j+99,j+148)]

        #combine for one df
        train.full <- (cbind(train.target,train.predictors))
        
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, c(j+1,j+50,j+99,j+148)]

        #find historical mean value
        historicalmeanpredictions.4[tau+1,j] <- mean(train.target)

        #find linear model prediction
        temp.mdl.linear <- lm(train.target ~ ., data=train.full)
        linearpredictions.4.own[tau+1,j] <- predict(temp.mdl.linear,
                                        newdata= tau.newdata)
        
        #find lasso model prediction
        temp.mdl.lasso.cv <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5)
        lassopredictionscv.4.own[tau+1,j] <- predict(temp.mdl.lasso.cv,
                                        newx = as.matrix(tau.newdata),
                                    type = "response")
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(linearpredictions.4.own,
     lassopredictionscv.4.own,
     file="ownpredictions4.RData")
```

```{r all4lagmodel}
set.seed(88)
lassopredictionscv.4.all <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
coefsum <- matrix(0, nrow = 49, ncol = (ncol(df.4)+1))

df <- df.4
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), 2:ncol(df)]

        #combine for one df
        train.full <- (cbind(train.target,train.predictors))
        
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, 2:ncol(df)]

        #find historical mean value
        historicalmeanpredictions.4[tau+1,j] <- mean(train.target)

        #find lasso model prediction
        temp.mdl.lasso.allcv <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5)
        
        #extract coefficients
        coefs <- as.matrix(coefficients(temp.mdl.lasso.allcv))
        coefsdf <- data.frame(variable =rownames(coefs),
                              coefficients=coefs[,1])
        
        #add 1 to the sum of coefficients if it is used in the model
        for (i in 2:nrow(coefsdf)) {
          if (coefsdf$coefficients[i] != 0) {
            print(paste("tau",tau,"coef",i,sep=""))
            coefsum[j,i] <- coefsum[j,i]+1}}
        
        lassopredictionscv.4.all[tau+1,j] <- predict(temp.mdl.lasso.allcv,
                                        newx = as.matrix(tau.newdata),
                                    type = "response")
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(historicalmeanpredictions.4,
     lassopredictionscv.4.all,
     coefsum,
     file="allpredictions4.RData")
```

```{r coef table}
coefcounttable <- matrix(0, nrow=49, ncol=8)
coefsum <- coefsum[,1:197]
for (i in 1:49){
  #assumption that this is without date
  coefcounttable[i,1] <- coefsum[i,i+1]
  coefcounttable[i,2] <- coefsum[i,i+50]
  coefcounttable[i,3] <- coefsum[i,i+99]
  coefcounttable[i,4] <- coefsum[i,i+148]
  
  temp.coefsdf <- data.frame(variable=coefsdf$variable,
                             sums= coefsum[i,])
  temp.coefsdf <- temp.coefsdf[!(1:nrow(temp.coefsdf)) %in% c(1,i+1,i+50, i+99,i+148),]

  #find the next other most used variable
  highestvar <- which.max(temp.coefsdf$sums)[1]
  if (temp.coefsdf$sums[highestvar] != 0) coefcounttable[i,5] <- as.character(temp.coefsdf$variable[highestvar])
  coefcounttable[i,6] <- temp.coefsdf$sums[highestvar]
  temp.coefsdf <- temp.coefsdf[!(1:nrow(temp.coefsdf)) %in% c(highestvar),]
  
  #find the second other most used variable
  highestvar <- which.max(temp.coefsdf$sums)[1]
  if (temp.coefsdf$sums[highestvar] != 0) coefcounttable[i,7] <- as.character(temp.coefsdf$variable[highestvar])
  coefcounttable[i,8] <- temp.coefsdf$sums[highestvar]
  temp.coefsdf <- temp.coefsdf[!(1:nrow(temp.coefsdf)) %in% c(highestvar),]
}
coefcounttable <- cbind(names(returns[2:ncol(returns)]),coefcounttable)

colnames(coefcounttable) <- c("Industry",
                              "Lag1","Lag2","Lag3","Lag4",
                              "V1", "V1 Usage",
                              "V2", "V2 Usage")
save(coefcounttable, file="coefcounttable.Rdata")
```

```{r all4lagmodeladvanced}
set.seed(88)
lagrfpredictions.4.all <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
laggbmpredictionsgauss.4.all <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
lagrfpredictions.4.cv.all <- matrix(0, ncol = 49, nrow = (nrow(df.4)+1))
rfmodelused <- vector()

df <- df.4
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), ]

        #combine for one df
        train.full <- (cbind(train.target,train.predictors))
        
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, ]
        
        #find random forest prediction with CV
        params <- list(c(40,2),c(50,2),c(50,3),c(50,4),c(80,2))
        error <- 100
        for (p in params)
        {
        cv.temp.mdl.randomforest <- randomForest(train.target ~ .,data=train.full,
                            ntree=p[1],
                            nodesize=p[2])
        if (cv.temp.mdl.randomforest$mse[p[1]] < error){
          temp.mdl.randomforest <- cv.temp.mdl.randomforest
          error <- cv.temp.mdl.randomforest$mse[p[1]]
          paramsused <- p}
        }
        rfmodelused <- rbind(rfmodelused,paramsused)
 
        lagrfpredictions.4.cv.all[tau+1,j] <- predict(temp.mdl.randomforest, tau.newdata)

         #find gaussian general boosting
        temp.mdl.gbm.gaus <- gbm(train.target ~ .,
                            data=train.full,
                            distribution="gaussian",
                            n.trees=50,
                            cv.folds=4)
        laggbmpredictionsgauss.4.all[tau+1,j] <- predict(temp.mdl.gbm.gaus,
                                              newdata=tau.newdata,
                                              n.trees=50,
                                              n.cores=1)

}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(lagrfpredictions.4.cv.all,
     laggbmpredictionsgauss.4.all,
     file="allpredictions4advancedcv.RData")
```

```{r lag4performanceprep}
lag4actualsdf <- returns[(windowlength+4):(nrow(returns)),2:ncol(returns)]

historicalmeanpredictions.4.clean <- as.data.frame(historicalmeanpredictions.4)[(windowlength+1):(nrow(historicalmeanpredictions.4)-1),]
historicalmean.4SE <- (historicalmeanpredictions.4.clean - lag4actualsdf)^2
historicalmean.4SSE <- colSums(historicalmean.4SE)

linearpredictions.4.own.clean <- as.data.frame(linearpredictions.4.own)[(windowlength+1):(nrow(linearpredictions.4.own)-1),]
linear.4.ownSE <- (linearpredictions.4.own.clean - lag4actualsdf)^2
linear.4.ownSSE <- colSums(linear.4.ownSE)

lassopredictionscv.4.own.clean <- as.data.frame(lassopredictionscv.4.own)[(windowlength+1):(nrow(lassopredictionscv.4.own)-1),]
lasso.cv.4.ownSE <- (lassopredictionscv.4.own.clean - lag4actualsdf)^2
lasso.cv.4.ownSSE <- colSums(lasso.cv.4.ownSE)

lassopredictionscv.4.all.clean <- as.data.frame(lassopredictionscv.4.all)[(windowlength+1):(nrow(lassopredictionscv.4.all)-1),]
lasso.cv.4.allSE <- (lassopredictionscv.4.all.clean - lag4actualsdf)^2
lasso.cv.4.allSSE <- colSums(lasso.cv.4.allSE)

lagrfpredictions.4.cv.all.clean <- 
as.data.frame(lagrfpredictions.4.cv.all)[(windowlength+1):(nrow(lagrfpredictions.4.cv.all)-1),]
lagrf.4.allcvlSE <- (lagrfpredictions.4.cv.all.clean - lag4actualsdf)^2
lagrf.4.allcvSSE <- colSums(lagrf.4.allcvlSE)

laggbmpredictionsgauss.4.all.clean <-
as.data.frame(laggbmpredictionsgauss.4.all)[(windowlength+1):(nrow(laggbmpredictionsgauss.4.all)-1),]
laggbmgauss.4.allSE <- (laggbmpredictionsgauss.4.all.clean - lag4actualsdf)^2
laggbmgauss.4.allSSE <- colSums(laggbmgauss.4.allSE)

save(lasso.cv.4.allSE,historicalmean.4SE, file="SEforGoyal.RData")
```

```{r 4lagnewsprep}
df.4.temp <- df.4
df.4.temp$date <- as.Date(as.character(df.4.temp$date), "%Y%m%d")
df.4.temp <- df.4.temp[df.4.temp$date <= "2016-07-01",]

#read in news data
industrynews <- read.csv("industrynews.csv", stringsAsFactors = FALSE)
#format date
industrynews$date <- as.Date(industrynews$date)
#remove X column
industrynews <- industrynews[,2:ncol(industrynews)]
#add _news to all news columns
for (i in 2:ncol(industrynews)) colnames(industrynews)[i] <- paste(colnames(industrynews)[i],"_news",sep="") 

df.4.news <- merge(df.4.temp,industrynews, by=c("date","date"))
```

```{r all4lagnewsmodel}
set.seed(88)
historicalmeanpredictions.4.news <- matrix(0, ncol = 49, nrow = (nrow(df.4.news)+1))
lassopredictionscv.4.news.all <- matrix(0, ncol = 49, nrow = (nrow(df.4.news)+1))
df <- df.4.news
for (tau in windowlength:nrow(df)) {
        for (j in 1:49){
        
        #creating target, train and prediction data
        #for window at tau, for industry j
        
        #take j values for time 2 onwards
        train.target <- df[(tau-windowlength+2):tau,j+1]
        
        #take away date column
        train.predictors <- df[(tau-windowlength+1):(tau-1), 2:ncol(df)]

        #combine for one df
        train.full <- (cbind(train.target,train.predictors))
        
        #newdata for predicting tau + 1
        tau.newdata <- df[tau, 2:ncol(df)]

        #find historical mean value
        historicalmeanpredictions.4.news[tau+1,j] <- mean(train.target)

        #find lasso model prediction
        temp.mdl.lasso.allcv <- cv.glmnet(as.matrix(train.predictors),
                               as.matrix(train.target),
                               alpha = 1, nfolds=5)
        lassopredictionscv.4.news.all[tau+1,j] <- predict(temp.mdl.lasso.allcv,
                                        newx = as.matrix(tau.newdata),
                                    type = "response")
}
##because this code takes a while to run, print for progress
print(paste(round((tau-80)/444,3)*100,"%",sep=""))
}
save(lassopredictionscv.4.news.all,
     historicalmeanpredictions.4.news,
     file="allpredictions4news.RData")
```

```{r lag4newsperformanceprep}
lagactuals4.news <- returns
lagactuals4.news$date <- as.Date(as.character(lagactuals4.news$date),"%Y%m%d")
lagactuals4.news <- lagactuals4.news[lagactuals4.news$date <= "2016-07-02",]
lagactuals4.news <- lagactuals4.news[(windowlength+4):(nrow(lagactuals4.news)),2:ncol(lagactuals4.news)]

historicalmeanpredictions.4.news.clean <- as.data.frame(historicalmeanpredictions.4.news)[(windowlength+1):(nrow(historicalmeanpredictions.4.news)-1),]
historicalmean.news.4SE <- (historicalmeanpredictions.4.news.clean - lagactuals4.news)^2
historicalmean.news.4SSE <- colSums(historicalmean.news.4SE)

lassopredictionscv.4.news.all.clean <- as.data.frame(lassopredictionscv.4.news.all)[(windowlength+1):(nrow(lassopredictionscv.4.news.all)-1),]
lasso.cv.news.4.allSE <- (lassopredictionscv.4.news.all.clean - lagactuals4.news)^2
lasso.cv.news.4.allSSE <- colSums(lasso.cv.news.4.allSE)
```

```{r r2oos}
R2oos <- data.frame(
                    linearown1 = round(1-(linear.1.ownSSE/historicalmeanSSE),4),
                    linearown4 = round(1-(linear.4.ownSSE/historicalmean.4SSE),4),
                    lassoown4 = round(1-(lasso.cv.4.ownSSE/historicalmean.4SSE),4),
                    linearall1 = round(1-(linear.1.allSSE/historicalmeanSSE),4),
                    lassall1 = round(1-(lasso.cv.1.allSSE/historicalmeanSSE),4),
                    lassoall4 = round(1-(lasso.cv.4.allSSE/historicalmean.4SSE),4),
                    lassoall4news = round(1-(lasso.cv.news.4.allSSE/historicalmean.news.4SSE),4),
                    rf4cv = round(1-(lagrf.4.allcvSSE/historicalmean.4SSE),4),
                    gauss4all = round(1-(laggbmgauss.4.allSSE/historicalmean.4SSE),4)
                    )
R2oos <- rbind(round(colMeans(R2oos),4),R2oos)
R2oos <- cbind(c("Average",colnames(returns[2:ncol(returns)])),R2oos)
names(R2oos) <- c("Industry","Lin Own 1", "Lin Own 4", "L Own 4",
                  "Lin 1", "L 1", "L 4",
                  "L 4 news", "RF 4", "GBM 4")
save(R2oos, file="R2oos.Rdata")
```

```{r condformattable}
##
#Note that it does not seem to be possible with the condtable package to run this in a for loop, or for all columns.
#http://stackoverflow.com/questions/42565960/conditional-formatting-while-looping-through-columns-using-package-condformat
##
condr2table <- condformat(R2oos) +
  rule_fill_discrete(2,
                     expression = R2oos[,2] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(3,
                     expression = R2oos[,3] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(4,
                     expression = R2oos[,4] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(5,
                     expression = R2oos[,5] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(6,
                     expression = R2oos[,6] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(7,
                     expression = R2oos[,7] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(8,
                     expression = R2oos[,8] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(9,
                     expression = R2oos[,9] > 0,
                     colours = c("TRUE" = "#c2f0c2")) +
  rule_fill_discrete(10,
                     expression = R2oos[,10] > 0,
                     colours = c("TRUE" = "#c2f0c2"))
save(condr2table,file="condR2table.RData")
```
